Scalar and Representative Models for Polarimetric SAR Data

### 2013-04-22 paper_dev/article_pp.tex
201303_Mar01
added 201304_Apr22

Over a real and captured image, while the paramaters $d$ and $L$ do not change for the whole image,
  the underlying $\Sigma_v$ is expected to differ from one region to the next.
Thus over an heterogeneous scene, the stochastic process for $|C_v|$ and $\ln |C_v|$ varies depending on the underlying signal $\Sigma_v$. 
In such context, Eqn. \ref{eqn:var_det_is_heteroskedastic} implies that the variance of $|C_v|$ also differs depending on the underlying signal $\Sigma_v$, which indicates its heteroskedastic property.
At the same time, in the log-transformed domain, Eqn. \ref{eqn:var_log_det_is_homoskedastic} shows that
  the variance of $\ln |C_v|$ is invariant and independent of $\Sigma_v$ manifesting its homoskedastic nature.

\section{Consistent Measures of Distance for POLSAR}
\label{sec:distance_measure}

Similar to the way dispersion and contrast is defined in our previous work \cite{Le_2013_TGRS_SAR_MSE},
  this section introduces the consistent sense of distance in a few different perspective.
First assumming that the true value of the underlying signal $\Sigma_v$ is known \textit{a priori},
the following random variables,
  namely ratio ($\mathbb{R}$) and log-distance ($\mathbb{L}$),
  are observable according to their definitions:
%Eqns. \ref{eqn:prod_chi_squared_rv} and \ref{eqn:sum_log_chi_squared_rv} lead straight to the definition of the following random variables, which is the :

\begin{eqnarray}
  \mathbb{R} &=& \frac{|C_v|}{|\Sigma_v|} \label{eqn:determinant_ratio_observables}\\
  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}
From another perspective where the POLSAR is known coming from an homogeneous area but the true value of the underlying signal $\Sigma_v$ is \textit{unknown}, consider the dispersion ($\mathbb{D}$) and contrast ($\mathbb{C}$) random variables being defined as:
\begin{eqnarray}
  \mathbb{D} &=& \ln{|C_v|} - avg(\ln{|C_v|}) \label{eqn:dispersion_observable}\\
  \mathbb{C} &=& \ln(|C_{v1}|) - \ln(|C_{v2}|) \label{eqn:contrast_observable}
\end{eqnarray}

Taking the results from Eqns. \ref{eqn:determinant_distribution}, \ref{eqn:log_determinant_distribution} and \ref{eqn:avg_log_det} we have
\begin{eqnarray}
\mathbb{R} &\sim& \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi^2 (2L-2i) \label{eqn:determinant_ratio_distribution} \\
\mathbb{L} &\sim&  \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln(2L)
\label{eqn:log_determinant_distance_distribution} \\ 
 \mathbb{D} &\sim& \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln{2} + k
\label{eqn:dispersion_distribution} \\ 
 \mathbb{C} &\sim& \sum^{d-1}_{i=0} \Delta(2L-2i)
\label{eqn:contrast_distribution}  
\end{eqnarray}
with $\Delta(2L) \sim \Lambda(2L) - \Lambda(2L)$
and $k=\sum^{d-1}_{i=0} \psi^0(L-i)$

@@

Since each elementary simulation component follows fixed distributions (i.e. $\chi^2(2L), \Lambda(2L), ... $),
  it is natural that these variables also follow fixed distributions.
Moreover, they are independent to $\Sigma_v$.
This serves as conclusive evidence that
  these random variables follows consistent and fixed distributions,
  regardless of the underlying signal $\Sigma_v$.

@@

This section describes an experiment to validate the models above against real-life captured data.
The validation procedure is quite straightforward.
Given that the stochastic models have been derived in the previous sections, 
  they can be graphically visualized by the histogram plots of the simulated data.
At the same time, the form of the real-life practical data is also observable via histogram plots of the data samples extracted from an homogeneous area.
Therefore, the theoretical models can be validated 
  if for the same parameters 
  the two plots match each other reasonably well.

As a demonstration, a homogeneous area was chosen from the AIRSAR Flevoland POLSAR data as experimental data samples.
Then theoretical models are employed in an attempt to explain the data.
The validating models include:
  the determinant and its log-transformed models, together with the dissimilarity measures namely: the determinant ratio, the log-distance, the dispersion and the contrast measures of distance.

The models are closely related.
Given the same parameter set, the determinant and determinant ratio are just scaled version of each other.
Meanwhile, the log-determinant, the log-distance and the dispersion are also just shifted version of each other.
Thus ideally speaking
  if one model is validated all the other models will also be,
  assuming that all the parameters of the image are known exactly.
Still in this section, all the models will be made subject to investigation.

Among all these models, the least-assumed stochastic process for dispersion and contrast measures of distance are validated first.
For each pixel in the region, the determinant of the covariance matrix is computed and the log-transformation is applied.
Then the average of POLSAR covariance matrix’s determinant in the log-transformed domain, i.e. $avg(ln|C_v|)$, is measured for dispersion.
Subsequently the observable samples of dispersion and contrast are computed according to Eqns. \ref{eqn:dispersion_observable} and \ref{eqn:contrast_observable}.
Then their histograms are traced out.

At the same time, theoretical simulations according to Eqns. \ref{eqn:dispersion_distribution} and \ref{eqn:contrast_distribution} are carried out.
The nominal value of 4 was taken as the dataset's number of look,
 while the dimensional number is set to 3 or 2 respectively depending on whether full or partial polarimetric SAR dataset is being investigated.
All the histogram plots are presented in Fig. \ref {fig:verify_polsar_2x2_simulation_dispersion_contrast}.
Apparently, a visual match is observable which verify the applicability of the theoretical models for the dispersion and contrast measures of distance.

\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[part-pol (2x2) dispersion]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_2x2_dispersion_distribution.eps} 
		 \label{dispersion_2x2}
	} 
	\hfill	
	\subfloat[part-pol (2x2) contrast]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_2x2_contrast_distribution.eps} 	
		 \label{contrast_2x2}
	} \\
	\subfloat[full-pol (3x3) dispersion]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_3x3_dispersion_distribution.eps} 
		 \label{dispersion_3x3}
	} 
	\hfill	
	\subfloat[full-pol (3x3) contrast]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_3x3_contrast_distribution.eps} 	
		 \label{contrast_3x3}
	}
\end{tabular}
\caption{Validating the dispersion and contrast models against both partial and full polarimetric AIRSAR Flevoland data.}
\label{fig:verify_polsar_2x2_simulation_dispersion_contrast}
\end{figure}

Apart from dispersion and contrast,
the other four models to be investigated require an estimation of the ``true'' underlying signal $|\Sigma_v|$. 
There are two ways to estimate this quantity over an homogeneous area.
The traditional way is to simply set the true signal equal to the average of the polsar covariance matrix in its original domain, i.e. $\Sigma_v = avg(C_v$).
Another approach is to estimate the true signal from the average of the log-determinant of the polsar covariance matrix (i.e. $avg[ln|C_v|]$) using Eqn. \ref{eqn:avg_log_det}.
Both approaches will be explored in this section.
However, given that the log-determinant average has already been computed earlier, 
  the second approach is used first for the validation of determinant-ratio and log-distance.

Fig. \ref{fig:verify_polsar_2x2_simulation_det_ratio_log_distance} validate the models of determinant-ratio and log-distance against real-life data.
In this experiment, the theoretical models is simulated from Eqns \ref{eqn:determinant_ratio_distribution} and \ref{eqn:log_determinant_distance_distribution},
  while the observable samples are computed using Eqns \ref{eqn:determinant_ratio_observables} and \ref{eqn:log_distance_observables}
  with the true signal estimated from the log-determinant average, i.e. $avg(\ln|C_v|)$.
Again a reasonable match is observed which validates the models for log-distance and determinant ratio.

\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[part-pol (2x2) determinant ratio]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_2x2_determinant_ratio_distribution.eps} 
		 \label{determinant_ratio_2x2}
	} 
	\hfill	
	\subfloat[part-pol (2x2) log distance]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_2x2_log_distance_distribution.eps} 	
		 \label{log_distance_2x2}
	} \\
	\subfloat[full-pol 3x3 determinant ratio]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_3x3_determinant_ratio_distribution.eps} 
		 \label{determinant_ratio_3x3}
	} 
	\hfill	
	\subfloat[full-pol 3x3 log distance]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_3x3_log_distance_distribution.eps} 	
		 \label{log_distance_3x3}
	}
\end{tabular}
\caption{Validating determinant-ratio and log-distance models with $|\Sigma_v|$ is computed using $avg(\ln|C_v|)$}
\label{fig:verify_polsar_2x2_simulation_det_ratio_log_distance}
\end{figure}

Since the models for the determinant and log-determinant are just scaled or shifted version of the models for determinant-ratio and log-distance, similar validation results are to be expected. 
And if the  true signal are computed in the same manner as described before then in fact similar match can be easily observed.

However, a more interesting  phenomena is to be described.
It happens in the validation process for determinant and its log-transformed model,
  where the theoretical response is taken from the simulated stochastic process described by Eqns. \ref{eqn:determinant_distribution} and \ref{eqn:log_determinant_distribution}.
%For such simulation, an estimation of  the true signal is needed.
The interesting phenomena happens when the true signal is estimated by the first approach i.e. equal to the average of the sample covariance matrix in its original domain.
Interestingly, this approach results in a different estimation for the true signal with reference to the method described earlier.

Subsequently the validation plots, which are presented in Fig \ref{fig:verify_polsar_2x2_simulation_det}, exhibit some small discrepancies.
The differences are easier to observe in the log-determinant plots, 
  where except for some small shifted effects, the shapes of the models and the practical data exhibits certain resemblance.

\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[verify polsar 2x2: determinant]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_2x2_determinant_distribution.eps} 
		 \label{determinant_2x2}
	} 
	\hfill	
	\subfloat[verify polsar 2x2: log-determinant]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_2x2_log_det_distribution.eps} 	
		 \label{log_det_2x2}
	} \\ 
	\subfloat[verify polsar 3x3: determinant]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_3x3_determinant_distribution.eps} 
		 \label{determinant_3x3}
	} 
	\hfill	
	\subfloat[verify polsar 3x3: log-determinant]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_3x3_log_det_distribution.eps} 	
		 \label{log_det_3x3}
	} 
\end{tabular}
\caption{Validating determinant and log-determinant models with $\Sigma_v = avg(C_v)$}
\label{fig:verify_polsar_2x2_simulation_det}
\end{figure}

In short the least-assumed dispersion and contrast measures of distance are shown to match reasonably well with the practical data.
The same can be stated for the other four models, namely: determinant, log-determinant, determinant ratio and log-distance,
  if the underlying parameters can be estimated reasonably well for the given image.   
However as described above a single ``true signal'' $|\Sigma_v|$ can have two different estimated values,
  depending on which estimation method was being used.
The discrepancy suggests that at least one parameter for the models was wrongly employed.

But what model parameter were used wrongly, and even if that can be corrected, would a better match become observable?
The question is answered in Section \ref{sec:improve_the_match_bw_theory_practice}, 
  where not only the look number is shown to be misused
  but also the match of between the theoretical model and the practical data is shown to improve as well once a better look number (ENL) is estimated.
For now, let us contend that:
  using appropriate estimation of the parameters, 
  the proposed models match reasonably well with the practical data.

>>>>

For a real world captured image, while the parameters $d$ and $L$ do not change for the whole image,
%Over a real and captured image, while the parameters $d$ and $L$ do not change for the whole image,
  the underlying $\Sigma_v$ is expected to differ from one region to the next.
Thus over an heterogeneous scene, the stochastic process for $|C_v|$ and $\ln |C_v|$ varies depending on the underlying signal $\Sigma_v$. 
In such context, Eqn. \ref{eqn:var_det_is_heteroskedastic} implies that the variance of $|C_v|$ also differs depending on the underlying signal $\Sigma_v$, which indicates its heteroskedastic property.
At the same time, in the log-transformed domain, Eqn. \ref{eqn:var_log_det_is_homoskedastic} reveals that
  the variance of $\ln |C_v|$ is invariant and independent of $\Sigma_v$ manifesting its homoskedastic nature.

\subsection{Consistent Measures of Distance for POLSAR}
\label{sec:distance_measure}

Similar to the way dispersion and contrast is defined in our previous work \cite{Le_2010_ACRS} \cite{Le_2013_TGRS_SAR_MSE},
  this section introduces the consistent sense of distance from a couple of different perspectives.
Assuming, on the one hand, that the true value of the underlying signal $\Sigma_v$ is known \textit{a priori},
random variables,
  ratio ($\mathbb{R}$) and log-distance ($\mathbb{L}$),
  are observable according to their definitions:
%Eqns. \ref{eqn:prod_chi_squared_rv} and \ref{eqn:sum_log_chi_squared_rv} lead straight to the definition of the following random variables, which is the :
\begin{eqnarray}
  \mathbb{R} &=& \frac{|C_v|}{|\Sigma_v|} \label{eqn:determinant_ratio_observables}\\
  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}

On the other hand, under a more forgiving assumption %From another perspective
  where the POLSAR is known to have come from a homogeneous area, but the true value of the underlying signal $\Sigma_v$ is \textit{unknown},
  the dispersion ($\mathbb{D}$) and contrast ($\mathbb{C}$) random variables are observable and they are defined as:
\begin{eqnarray}
  \mathbb{D} &=& \ln{|C_v|} - avg(\ln{|C_v|}) \label{eqn:dispersion_observable}\\
  \mathbb{C} &=& \ln(|C_{v1}|) - \ln(|C_{v2}|) \label{eqn:contrast_observable}
\end{eqnarray}

Using the results from Eqns. \ref{eqn:determinant_distribution}, \ref{eqn:log_determinant_distribution} and \ref{eqn:avg_log_det} we have
\begin{eqnarray}
\mathbb{R} &\sim& \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi^2 (2L-2i) \label{eqn:determinant_ratio_distribution} \\
\mathbb{L} &\sim&  \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln(2L)
\label{eqn:log_determinant_distance_distribution} \\ 
 \mathbb{D} &\sim& \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln{2} + k
\label{eqn:dispersion_distribution} \\ 
 \mathbb{C} &\sim& \sum^{d-1}_{i=0} \Delta(2L-2i)
\label{eqn:contrast_distribution}  
\end{eqnarray}
with $\Delta(2L) \sim \Lambda(2L) - \Lambda(2L)$
and $k=\sum^{d-1}_{i=0} \psi^0(L-i)$

@@

Since each elementary component follows fixed distributions (i.e. $\chi^2(2L), \Lambda(2L), ... $),
  it is natural that these variables also follow fixed distributions.
Moreover, they are independent of $\Sigma_v$.
%This result shows how
In short, these random variables are shown to follow consistent and fixed distributions,
  regardless of the underlying signal $\Sigma_v$.

The previous section has introduced the theoretical model for 3-dimensional $d=3$ full polarimetric and two dimensional $d=2$ partial polarimetry cases.
In this section, the model is shown to be also applicable for the 1-dimensional $d=1$ case.
Physically this means the multi-dimensional POLSAR dataset is collapsed into one-dimensional conventional SAR data.
Mathematically, the sample covariance matrix is reduced to the sample variance
%The prevous section has validated the use of our models for 3-dimensional $d=3$ full polarimetric and two dimensional $d=2$ partial polarimetry cases.
%In this section, the focus is on the case where the dimensional number is reduced to 1 $d=1$.
%Physically this means the multi-dimensional POLSAR dataset is collapsed into the one-dimensional and classical SAR data.
%Mathmatically, the sample covariance matrix is reduced to the sample variance
  and the determinant equates the scalar value.
On the other hand, it is well known that for SAR data, variance equals intensity.
Thus the special case of our result is investigated carefully and is shown to be consistent with previous results for SAR intensity data.
This can be thought of 
  either as a cross-validation evidence for the proposed POLSAR models
  or alternatively as having SAR as the special case of POLSAR. 
%  with well known results for SAR intensity is presented.

%The results so far for our models can be summarized as:
The results so far for our models can be summarized using the following equations:
\begin{align}
  \mathbb{R} &= \frac{|C_v|}{|\Sigma_v|} \sim \frac{1}{(2L)^d} \prod^{d-1}_{i=0} \chi^2(2L-2i) \\% \label{eqn:polsar_ratio_det_cov_dist} \\
  \mathbb{L} &= \ln{|C_v|} - \ln{|\Sigma_v|} \sim \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \cdot \ln{2L} \\ %\label{eqn:polsar_dispersion_log_det_cov_dist} \\
  \mathbb{D} &= \ln{|C_v|} - avg(\ln{|C_v|}) \sim \sum^{d-1}_{i=0} \Lambda(2L-2i) - d \ln{2} + k\\
  \mathbb{C} &= \ln{|C_{1v}|} - \ln{|C_{2v}|} \sim \sum^{d-1}_{i=0} \Delta(2L-2i) \\
  \mathbb{A} &= avg(\mathbb{L}) = \sum^{d-1}_{i=0} \psi^0(L-i) - d \cdot \ln{L} \\ %\label{eqn:polsar_dispersion_averages} \\
  \mathbb{V} &= var(\mathbb{L}) = \sum^{d-1}_{i=0} \psi^1(L-i) \\ %\label{eqn:polsar_dispersion_variance} \\
  \mathbb{E} &= mse(\mathbb{L}) =\left[ \sum^{d-1}_{i=0} \psi^0(L-i) - d \cdot \ln{L} \right]^2 +  \sum^{d-1}_{i=0} \psi^1(L-i) \label{eqn:polsar_dispersion_mse} 
\end{align}

Upon setting $d=1$ into the above equations,
  Appendix \ref{sec:appendix_sar_special_case_of_polsar} shows that the reduced results are consistent with the following two cases.
First is the the following results obtained from the our previous works on single-look SAR \cite{Le_2010_ACRS} \cite{Le_2013_TGRS_SAR_MSE}, i.e. $d=L=1$,
%Upon setting $d=1$ into the above models,
%  Appendix \ref{sec:appendix_sar_special_case_of_polsar} shows that the reduced results are consistent with
%not only the following results from the our previous works on single-look SAR \cite{Le_2013_TGRS_SAR_MSE}, i.e. $d=L=1$,
\begin{align*}
  I &\sim \bar{I} \cdot pdf \left[ e^{-R} \right] \\
  \log_2{I} &\sim \log_2{\bar{I}} + pdf \left[ 2^xe^{-2^x}\ln2 \right] \\
  \mathbb{R} &= \frac{I}{\bar{I}} \sim pdf \left[ e^{-x} \right]  \\
  \mathbb{L} &= \log_2{I} - \log_2{\bar{I}} \sim pdf \left[ 2^xe^{-2^x}\ln2 \right]\\
  \mathbb{D} &= \log_2{I} - avg(\log_2{I}) \sim pdf \left[ e^{-(2^xe^{-\gamma})} 2^xe^{-\gamma} \ln2 \right] \\
  \mathbb{C} &= \log_2{I_1} - \log_2{I_2} \sim pdf \left[ \frac{2^x}{(1+2^x)^2} \ln2 \right] \\
  \mathbb{A} &= avg(\mathbb{L}) = -\gamma / \ln{2} \\
  \mathbb{V} &= var(\mathbb{L}) = \frac{\pi^2}{6} \frac{1}{ \ln^2{2}} \\
  \mathbb{E} &= mse(\mathbb{L}) = \frac{1}{\ln^2{2}}( \gamma^2 + \pi^2/6 ) = 4.1161 
\end{align*}
The second is the following well-known results for multi-look SAR, i.e. $d=1,L>1$:
%but also the following well-known results for multi-look SAR, i.e. $d=1,L>1$:
  \begin{eqnarray}
I &\sim& pdf \left[ \frac{L^L x^{L-1} e^{-Lx/\bar{I}}}{\Gamma(L) \bar{I}^L} \right] \\
N = \ln{I} &\sim& pdf \left[ \frac{L^L}{\Gamma(L)} e^{L(x-\bar{N})-Le^{x-\bar{N}}} \right]
  \end{eqnarray}
Furthermore, the following derivations for multi-look SAR data
   show that it can be thought of 
    either as extensions of the corresponding single-look SAR results
    or as simple cases of the POLSAR results
  are also derived as:
%  Furthermore, the following results for multi-look SAR data,
%  which can be thought of 
%    either as extensions of the corresponding single-look SAR results
%    or as simple cases of the POLSAR results
%  are also derived as:
  \begin{align*}
    \mathbb{R} &= \frac{I}{\bar{I}} \sim pdf \left[ \frac{ L^{L} x^{L-1} e^{-Lx}}{ \Gamma(L)} \label{eqn:multi_look_SAR_ratio_dist} \right]\\
    \mathbb{L} &= \ln{I} - \ln{\bar{I}} \sim pdf \left[ \frac{L^Le^{Lt-Le^t}}{ \Gamma(L)}  \right] \\
    \mathbb{D} &= \ln{I} - avg(\ln{I}) \sim pdf \left[ \frac{e^{L[x-\psi^0(L)]-e^{[x-\psi^0(L)]}}}{\Gamma(L)} \right] \\
    \mathbb{C} &= \ln{I_1} - \ln{I_2} \sim pdf \left[ \frac{e^{x}}{(1+e^x)^{2}} \right] \\
    \mathbb{A} &= avg(\mathbb{L}) = \psi^0(L) - \ln{L} \\
    \mathbb{V} &= var(\mathbb{L}) = \psi^1(L) \\
    \mathbb{E} &= mse(\mathbb{L}) = \left[ \psi^0(L) - \ln{L} \right]^2 + \psi^1(L)
  \end{align*}

This newly derived models for multi-look SAR data can also be validated against real-life data.
Fig. \ref{fig:verify_multi_look_SAR_dispersion_contrast_models} presents the results of an experiment carried out for the stated purpose.
In the experiment the intensity of a single-channel SAR data (HH) for a homogeneous area in the AIRSAR Flevoland dataset is extracted.
The histograms for the log-distance and contrast is then plotted against the theoretical PDF given above.
 The plot is obtained with ENL set to the nominal number of 4, and  good visual match is apparent in the final results.
%Then the histograms for the log-distance and and contrast is plotted against the theoretical PDF given above.
%The ENL is set to the nominal number of 4.
%And good visual match is apparent in the final results.

%ifdef FINAL
\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[multi-look SAR dispersion]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_multi_look_sar_dispersion_pdf.eps} 	
		 \label{multi_look_dispersion}
	} 
	\hfill	
	\subfloat[multi-look SAR contrast]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_multi_look_sar_contrast_pdf.eps} 	
		 \label{multi_look_contrast}
	}
\end{tabular}
\caption{Multi-Look SAR dispersion and contrast: modelled response matches very well with real-life captured data.}
\label{fig:verify_multi_look_SAR_dispersion_contrast_models}
\end{figure}
%endif
%ifdef REVIEW
\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[multi-look SAR log-distance]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/verify_multi_look_sar_dispersion_pdf.eps} 	
		 \label{multi_look_dispersion}
	} 
	\hfill	
	\subfloat[multi-look SAR contrast]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/verify_multi_look_sar_contrast_pdf.eps} 	
		 \label{multi_look_contrast}
	}
\end{tabular}
\caption{Multi-Look SAR dispersion and contrast: modelled response matches very well with real-life captured data.}
\label{fig:verify_multi_look_SAR_dispersion_contrast_models}
\end{figure}
%endif

@@

This section validates the theoretical models developed above against real-life practical data.
The first sub-section provides a naive validation, where the nominal ENL is used for the model.
While the match appears to be reasonably good, it can be further improved.
After the second sub-section discusses why the nominal look-number given by (POL)SAR processors may not be very accurate,
  the third sub-section proposes a new technique in estimating the Effective Number of Looks (ENL) using the consistent variance found in homoskedastic model.
With the newly estimated ENL, the final sub-section shows that the match between the theoretical model and the practical data is indeed further improved.

\subsection{Using the nominal ENL to validate the theoretical models} 
This sub-section describes an experiment to validate the models presented earlier against real-life captured data,
  which can be performed in a rather straightforward manner as follows.
The stochastic models derived in the previous sections can be graphically visualized as histogram plots of the simulated data.
%This section describes an experiment to validate the models above against real-life captured data.
%The validation procedure is quite straightforward.
%Given that the stochastic models have been derived in the previous sections, 
%  they can be graphically visualized by the histogram plots of the simulated data.
At the same time, the form of the real-life practical data is also observable via the histogram plots of the data samples extracted from an homogeneous area.
Therefore, the theoretical models can be validated 
  if for the same parameters 
  the two plots match each other reasonably well.

For this purpose, a homogeneous area was chosen from the AIRSAR Flevoland POLSAR data as experimental data samples.
Then theoretical models ar used  to explain the data.
%As a demonstration, a homogeneous area was chosen from the AIRSAR Flevoland POLSAR data as experimental data samples.
%Then theoretical models are employed in an attempt to explain the data.
The validating models include:
  the determinant and its log-transformed models, together with the dissimilarity measures namely: the determinant ratio, log-distance, dispersion and the contrast measures of distance.

These models are closely related as follows.
For  the same parameter set, the determinant and determinant ratio are simply scaled versions of each other.
Meanwhile, the log-determinant, log-distance and dispersion are also just shifted versions of each other.
Hence one could expect that 
  once a model is validated, the other models in the set will follow suit, 
  assuming that all the parameters of the image are known.% exactly.
Nevertheless, all the models will be separately evaluated in this section.
%The models are closely related.
%Given the same parameter set, the determinant and determinant ratio are just scaled version of each other.
%Meanwhile, the log-determinant, the log-distance and the dispersion are also just shifted version of each other.
%Thus ideally speaking
%  if one model is validated all the other models will also be,
%  assuming that all the parameters of the image are known exactly.
%Still in this section, all the models will be made subject to investigation.

Among all these models, the least-assumed stochastic process for dispersion and contrast measures of distance are validated first.
For each pixel in the region, the determinant of the covariance matrix is computed and the log-transformation is applied.
Then the average of POLSAR covariance matrix’s determinant in the log-transformed domain, i.e. $avg(ln|C_v|)$, is measured for dispersion.
Subsequently the observable samples of dispersion and contrast are computed according to Eqns. \ref{eqn:dispersion_observable} and \ref{eqn:contrast_observable} in order to plot their histograms.

Similarly, theoretical simulations according to Eqns. \ref{eqn:dispersion_distribution} and \ref{eqn:contrast_distribution} are carried out. $L=4$ for this study, while the dimensional number is set to either 3 or 2 depending on whether a full or partial polarimetric SAR dataset is being investigated.
The plots are presented in Fig. 
\ref {fig:verify_polsar_2x2_simulation_dispersion_contrast} showing an evident visual match between the model and real data, thus effectively validating the theoretical models for dispersion and contrast measures of distance.

%ifdef FINAL
\begin{figure}[h]
\centering
\begin{tabular}{c}
	\subfloat[part-pol (2x2) dispersion]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_2x2_dispersion_distribution.eps} 
		 \label{dispersion_2x2}
	} 
	\hfill	
	\subfloat[part-pol (2x2) contrast]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_2x2_contrast_distribution.eps} 	
		 \label{contrast_2x2}
	} \\
	\subfloat[full-pol (3x3) dispersion]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
                 \epsffile{images/verify_polsar_3x3_dispersion_distribution.eps} 
		 \label{dispersion_3x3}
	} 
	\hfill	
	\subfloat[full-pol (3x3) contrast]{
		 \epsfxsize=1.5in
		 \epsfysize=1.5in
		 \epsffile{images/verify_polsar_3x3_contrast_distribution.eps} 	
		 \label{contrast_3x3}
	}
\end{tabular}
\caption{Validating the dispersion and contrast models against both partial and full polarimetric AIRSAR Flevoland data.}
\label{fig:verify_polsar_2x2_simulation_dispersion_contrast}
\end{figure}
%endif
%ifdef REVIEW
\begin{figure}[h!]
\centering
\begin{tabular}{c}
	\subfloat[part-pol (2x2) dispersion]{
		 \epsfxsize=3in
		 \epsfysize=3in
                 \epsffile{images/verify_polsar_2x2_dispersion_distribution.eps} 
		 \label{dispersion_2x2}
	} 
	\hfill	
	\subfloat[part-pol (2x2) contrast]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/verify_polsar_2x2_contrast_distribution.eps} 	
		 \label{contrast_2x2}
	} \\
	\subfloat[full-pol (3x3) dispersion]{
		 \epsfxsize=3in
		 \epsfysize=3in
                 \epsffile{images/verify_polsar_3x3_dispersion_distribution.eps} 
		 \label{dispersion_3x3}
	} 
	\hfill	
	\subfloat[full-pol (3x3) contrast]{
		 \epsfxsize=3in
		 \epsfysize=3in
		 \epsffile{images/verify_polsar_3x3_contrast_distribution.eps} 	
		 \label{contrast_3x3}
	}
\end{tabular}
\caption{Validating the dispersion and contrast models against both partial and full polarimetric AIRSAR Flevoland data.}
\label{fig:verify_polsar_2x2_simulation_dispersion_contrast}
\end{figure}
%endif

Apart from dispersion and contrast,
the other four models to be investigated require an estimation of the ``true'' underlying signal $|\Sigma_v|$. 
There are two ways to estimate this quantity over an homogeneous area.
The traditional way is to simply set the true signal equal to the average of the POLSAR covariance matrix in its original domain, i.e. $\Sigma_v = avg(C_v$).
Another approach is to estimate the true signal from the average of the log-determinant of the POLSAR covariance matrix (i.e. $avg[ln|C_v|]$) using Eqn. \ref{eqn:avg_log_det}.
Both approaches will be presented in this section.
As the log-determinant average has already been computed earlier, 
  the second approach is hence used first for the validation of determinant-ratio and log-distance.
%Both approaches will be explored in this section.
%However, given that the log-determinant average has already been computed earlier, 
%  the second approach is used first for the validation of determinant-ratio and log-distance.

Fig. \ref{fig:verify_polsar_2x2_simulation_det_ratio_log_distance} plots the determinant-ratio and log-distance models against real-life data.
In this experiment, the theoretical models are simulated using Eqns \ref{eqn:determinant_ratio_distribution} and
%Fig. \ref{fig:verify_polsar_2x2_simulation_det_ratio_log_distance} validate the models of determinant-ratio and log-distance against real-life data.
%In this experiment, the theoretical models is simulated from Eqns \ref{eqn:determinant_ratio_distribution}
and \ref{eqn:log_determinant_distance_distribution},
  while the observable samples are computed using Eqns \ref{eqn:determinant_ratio_observables} and \ref{eqn:log_distance_observables}
  with the true signal estimated from the log-determinant average, i.e. $avg(\ln|C_v|)$.
%Again a reasonable match is observed which validates the models for log-distance and determinant ratio.  
A reasonable match is again observed which validates the models for log-distance and determinant ratio.
